---
title: "Assignment2"
author: "Harley Combest"
date: "`r Sys.Date()`"
output: 
    html_document:
      toc: yes
      toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# 17/17


## Question 1

### (a)

$$P(ExpertFail|MatchedPrints)=(100-92.12)\%=7.88\%$$

### (b)

$$P(NoviceFail|MatchedPrints)=(100-74.55)\%=25.45\%$$

### (c)

The participant is more likely to be a novice.

## Question 2

### (a)

$$P(Positive|User)=50/100=.5$$

### (b)

$$P(Negative|Nonuser)=891/900=.99$$

### (c)

$$P(User|Positive)=\frac{P(Positive|User)P(User)}{P(Positive|User)P(User)+P(Positive|Nonuser)P(Nonuser)}=\frac{(.5)(.1)}{(.5)(.1)+(.01)(.99)}=0.8347245$$

## Question 3

Proof (Multiplicative Rule).

We want to prove that the number of different samples that can be formed from k sets of elements (where n represents individual sets of elements) is equivalent to
$$n_1n_2n_3\dots n_k. $$
We have that the integers is an integral domain and, therein, has the multiplication operation such that for any integer n or m we have
$$n*m=+^nm=m+m+\dots +m$$
where m +...+ m represents n copies of m.


Base Case-

Suppose we have 
$$n_1n_2.$$

We have by the multiplication operation that this is equivalent to k=2 copies of n.


Inductive Case-

Let's if this works for k+1

$$n_1n_2\dots n_{k+1}$$

By the multiplication operation, we have that this is equivalent to k+1 copies of n.

Therefore, the number of different samples that can be formed from k sets of elements (where n represents individual sets of elements) is equivalent to
$$n_1n_2n_3\dots n_k. $$

Q.E.D.

## Question 4

Proof(Permutations Rule).

We want to prove that for a set of N distinct elements that the number of arrangements one can place n elements from N into n different positions is equal to 

$$N(N-1)(N-2)\dots (N-n+1)=\frac{N!}{(N-n)!_.}$$

We know from the previous theorem that the number of different samples we can form from a number of sets between sets is the product between the number of sets and the individual sets.

Now for this proof, we want to find the number of arrangments of elements is 'distinct' order.


Base Case-

This means that for N ways of filling the first position we have N-1 ways of filling the second position.


Inductive Case- 

Suppose we want to fill up the n+1th space.

We have
$$N-(n+1)+1 = N-n$$
ways of filling that space.

Therefore, for a set of N distinct elements the number of arrangements one can place n elements from N into n different positions is equal to 

$$N(N-1)(N-2)\dots (N-n+1)=\frac{N!}{(N-n)!_.}$$
Q.E.D.

## Question 5

Proof (Partitions Rule).

The number of ways on can partition a set of N elements into k sets is equal to

$$\frac{N!}{n_1!n_2!\dots n_k}$$
where
$$n_1+n_2+\dots + n_k=N.$$

Suppose we arrange N distinct elements into N positions.

By The Permutations Rule, we have that there are

$$P^N_N=\frac{N}{(N-N)!}=N!$$
ways of doing this.

It follows that there are
$$n_1!$$

ways of permuting a group with 
$$n_1$$

elements.

Suppose that 
$$N=n_1+n_2+\dots+n_k.$$

We then have that the number of ways to partition N is equal to

$$\frac{N!}{n_1!n_2!\dots n_k!}$$

as desired.

Q.E.D.

## Question 6

Proof(Combinations Rule).

We want to show that the number of different samples of n elements that can be selected from N is equal to

$$\frac{N!}{n!(N-n)!}.$$

Unlike with permutations, the order in which the n elements are drawn doesn't matter.

From The Permutations Rule, we have that the number ways we can arrange n elements from a set of N elements in a distinct order is equal to
$$\frac{N!}{(N-n)!_.}$$

Notice that what we are trying to prove is essentially a partition into two groups, one of which is (N-n) the number of elements not selected for the sample, is given.

We have that the other group is n elements selected from the sample.

By The Partitions Rule, we then have

$$\frac{N!}{n!(N-n)!}$$

as desired.

Q.E.D.

## Question 7

### (a)

$$.09+.30+.37+.20+.04=.39+.37+.24=.39+.61=1.00$$

### (b)

$$P(3\cup 4)=.20+.04=.24$$

### (c)

$$P(0\cup 1)=.09+.30=.39$$

## Question 8

### a

Requirements for a Discrete Probability Distribution
$$(1.)\; 0\leq p(y) \leq 1$$
$$(2.)\sum_{all \; y}p(y)=1$$

Each p(y) satisfies (1.).

$$.17+.10+.11+.11+.10+.10+.07+.05+.03+.02+.02+.02+.02+.02+.01+.01+.01+.01+.01+.005+.005\\ = .27+.42+.15+.10+.05+.01\\=.69+.31=1.00$$

The total p(y) satisfies (2.).

### (b)

$$P(Y\geq 10)= P(Y=10) +\dots+P(Y=20)\\=.02+.02+.02+.02+.01+.01+.01+.01+.01+.005+.005\\=.08+.05+.01=.14$$

### (c)

$$E(Y)=\sum_{all\; y}yp(y)=0(.17)+\dots +20(.005)\\=0+\dots +.1\\=.32+\dots +.195\\=1.05+ \dots +.545\\=2.56+1.55+.545\\=2.56+2.095\\=4.655$$

$$V(Y)=E(Y^2)-E(Y)^2\\E(Y^2)=0^2(.17)+\dots+20^2(.005)=41.525\; and\;E(Y)^2=4.655*4.655=21.669025\\V(Y)=41.525-21.669025=19.855975$$

### (d)

$$P(?\leq Y \leq \;?)\geq.75\\P(0\leq Y\leq 6) = .76$$
Tbh, I was thinking about just putting down the entire interval as that would fulfill the prompt but I get the feeling that ain't what we're looking for.

## Question 9

### (a)

```{r}
x=choose(25,10)*(.7)^10*(.3)^15
x
```


$$P(Y=10)= {n\choose y}p^yq^{n-y}={25\choose 10}(.70)^{10}(.30)^{25-10}\approx 0.0013$$

### (b)

$$P(Y\leq 5)=P(Y=5)+\dots+P(Y=0)\\={25\choose 5}(.70)^{5}(.30)^{25-5}+\dots+ {25\choose0}(.70)^0(.30)^{25-0}\\ \approx3.457*10^{-7}$$

### (c)

$$\mu = np = (25)(.7) = 17.5\\ \delta=\sqrt{npq} =\sqrt{(25)(.7)(.3)}\approx2.2913$$

### (d)

On average, a random sample of 25 engineering students who recently earned their PhD will have roughly 17.5 foreign nationals.

There won't be too much variation in this as the standard deviation is low and indicates that the values of the data are relatively close to the mean.

## Question 10

### (a)
$$P(5, 5, 5, 5, 5, 5, 5, 5, 5, 5)= \frac{50!}{5!5!5!5!5!5!5!5!5!5!}(.1)^{5 *10} \approx 4.9121*10^{43}$$

### (b)

$$P(Y = 1) = \frac{50!}{1!49!}(.02)^{1}(.98)^{49}\approx.3716$$

## Question 11

### (a)

The Geometric Probability Distribution with Y = Number of trials until first success is found.

$$p(y)= pq^{y-1} = (.4)((45+15)/100)^{y-1} = (.4)(.6)^{y-1}$$

### (b)

$$E(Y) = \frac{1}{p} = \frac{1}{.4} = 2.5$$

On average, it'll take roughly 2.5 consumers who must be interviewed until one indicates something other than information given directly on the product's label or packing as the reason a product is green.

### (c)

$$P(Y=1)=p(1)=(.4)(.6)^{1-1}=.4$$

### (d)

$$P(Y>2)= 1 -P(Y=2)-P(Y=1)=1-[(.4)(.6)^{2-1}]-.4=.36$$

## Question 12

### (a)

$$\mu = \frac{nr}{N} = \frac{(10)(8)}{209}\approx .3828$$
In an average sample of 10 facilities, roughly .3828 of them treat hazardous waste on sight.


### (b)

$$P(Y=4)=\frac{{r\choose y}{N-r\choose n-y}}{{N\choose n}}=\frac{{8\choose 4}{209-8\choose 10-4}}{{209\choose 10}}\approx1.6885*10^{-4}$$

## Question 13

### (a)

$$\delta^2=\lambda=.03$$

### (b)

The number of casualties that occur in one deep-draft U.S. flag vessel over a 3-year period is independent of the number that occur in other deep-draft U.S. flag vessels.

Another condition, is that the probability of a casualty in a given deep-draft U.S. flag vessel over a 3-year period is the same for all other deep-draft U.S. flag vessels.

### (c)

$$P(Y=0)=p(0)=\frac{\lambda^0 e^{-\lambda}}{0!}=e^{-.03}\approx.9704$$

## Question 14

### (a)

$$\int_{-\infty}^{\infty}f(y)dy=\int_{0}^{1}c(2-y)dy=c[2y-\frac{1}{2}y^2]_{0}^{1} =\frac{3}{2}c=1 \implies c=\frac{2}{3}$$

### (b)

$$F(y)=\int_{-\infty}^{y}f(t)dt = \int_{0}^{y}(\frac{4}{3}-\frac{2}{3}t)dt=[\frac{4}{3}t-\frac{1}{3}t^2]_{0}^{y}=\frac{1}{3}(4y-y^2)$$

### (c)

$$F(.4)=\frac{1}{3}[4(.4)-(.4)^2]=.48$$

### (d)

$$P(.1\leq Y \leq .6)=\int_{.1}^{.6} f(y) dy=\int_{.1}^{.6}(\frac{4}{3}-\frac{2}{3}y^2)dy=[\frac{4}{3}y-\frac{1}{3}y^2]_{.1}^{.6}=[\frac{4}{3}(.6)-\frac{1}{3}(.6)^2] - [\frac{4}{3}(.1)-\frac{1}{3}(.1)^2] = .55$$

## Question 15

### (a)

Mean
$$E(Y)=\int_{-\infty}^{\infty}yf(y)dy=\frac{3}{500}\int_{-5}^{5}(25y-y^3)dy=\frac{3}{500}[\frac{25}{2}y^2-\frac{1}{3}y^3]_{-5}^{+5}=-.5$$

Variance
$$\delta^2=E(Y^2)-E(Y)^2\\E(Y^2)=\int_{-\infty}^{\infty}y^2f(y)dy=\frac{3}{500}\int_{-5}^{5}(25y^2-y^3)dy=\frac{3}{500}[\frac{25}{3}y^3-\frac{1}{4}y^4]_{-5}^{+5}=\frac{25}{2}\\\delta^2 = \frac{25}{2}-(\frac{-1}{2})^2=\frac{49}{4}$$

### (b)

Mean in hours by Theorem 5.1
$$E(\frac{1}{60}Y) =\frac{1}{60}E(Y)=-\frac{1}{120}$$

Variance in hours by Theorem 5.3
$$V(cY)=\delta^2_{cY}=c^2\delta^2=(\frac{1}{60})(\frac{49}{4})=\frac{49}{240}$$

### (c)

Mean
$$60E(Y)=-30$$


Variance
$$60^2\delta^2=360*\frac{49}{9}=1960$$

## Question 16

### (a)
Mean = 50, Standard Deviation =3.2
$$P(Y>45)=P(Y\geq46)\\Z=\frac{Y-\mu}{\delta}=\frac{46-50}{3.2}=-1.25\\P(Y\geq46)=P(Z\geq-1.25)=1-P(Z<-1.25)=1-.10565\approx.8944\\$$

### (b)

$$P(Y<55)\\Z=\frac{55-50}{3.2}=1.5625\\P(Y<55)=P(Z<1.5625)\approx.9409$$

### (c)

$$P(51<Y<52)=P(Y < 52) - P(Y<51)\\Z_1=\frac{52-50}{3.2}=.625\\Z_2=\frac{51-50}{3.2}=.3125\\P(Y < 52) - P(Y<51)=P(Z<.625)-P(Z<.3125)\approx.7340-.6229=.1111$$

## Question 17

### (a)
Mean = 605, Standard Deviation = 185
$$P(500<Y<700)=P(Y < 700) - P(Y<500)\\Z_1=\frac{700-605}{185}\approx.5135\\Z_2=\frac{500-605}{185}\approx-.5676\\P(Y < 700) - P(Y<500)\approx P(Z<.5135)-P(Z<-.5676)\approx.6960-.2850=.4110$$

### (b)

$$P(400<Y<500)=P(Y < 500) - P(Y<400)\\Z_1=\frac{500-605}{185}\approx-.5676\\Z_2=\frac{400-605}{185}\approx-1.1081\\P(Y < 500) - P(Y<400)\approx P(Z<-.5676)-P(Z<-1.1081)\approx.2850-.1339=.1511$$

### (c)

$$P(Y<850)=\\Z=\frac{850-605}{185}\approx1.3243\\P(Y < 850)\approx P(Z<1.3243)\approx.9073$$

### (d)

$$P(Y>1000)=\\Z=\frac{1000-605}{185}\approx2.1351\\P(Y >1000)\approx P(Z > 2.1351)= 1 - P(Z\leq 2.1351)\approx 1 -.9836 = .0164$$

### (e)

$$P(Z>\;?)=1-P(Z< \;?) =.10 = 1 - P(Z<1.28)=P(Z>1.28)\\Z=1.28=\frac{?\;-605}{185}\\?=(1.28*185)+605=841.8$$