---
title: "assignment3"
author: "Harley Combest"
date: "`r Sys.Date()`"
output: 
    html_document:
      toc: yes
      toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 13/16

## Question 1 (MS 5.54 - pg 215)

### a.

```{r}
prob1=exp(-120/95)
prob1
```

### b.

```{r}
fish=read.csv("PHISHING.csv")
mean(fish$INTTIME)#B=95ish
sd(fish$INTTIME)#B=91ish
```

The data almost follows an exponential distribution with beta equal to 95 but it's a bit off in the mean and variance. There might be some collaboration or collusion from within the company.

## Question 2 (MS 5.56 - Page 215)

### a.

```{r}
mean1=(3)*(.07)#mean=.21
mean1
var1=(3)*(.07)^2#variance=.0147
var1
```

### b.

```{r}
mean1+3*(sqrt(var1))
mean1-3*(sqrt(var1))
```

I wouldn't expect that and that seems to be an outlier as that value is beyond 3 standard deviations of the mean.

## Question 3 (MS 5.60 - Page 216)

### a.

```{r}
mean2=(2)*(2)#mean for A
mean2
mean3=(1)*(4)#mean for B
mean3
```

The mean for both formulai is 4.

### b.

```{r}
var2=(2)*(2)^2
var2
var3=(1)*(4)^2
var3
```

var(formula A)=8 and var(formula B)=16.

### c.

```{r}
fctn1 <- function(y){
  (1/4)*y*exp(-y/2)
}
prob2=integrate(fctn1, lower=0, upper=1)
prob2#formula A

fctn2 <- function(y){
  (1/4)*exp(-y/4)
}
prob3=integrate(fctn2, lower=0, upper=1)
prob3#formula B
```

P(Y<1|B)>P(Y<1|A):formula B has higher probability of generating a human reaction in less than 1 minute.

## Question 4 (MS 5.74 - Page 219)

### a. 

```{r}
fctn3<-function(x){(1/2)*x*exp((-x^2)/4)}
prob4=integrate(fctn3,lower=0,upper=2)
prob4
```

### b.

```{r}
mean4=(4)^(1/2)*.88623#obtained from table 6 appendix B
mean4
var4=(4)*(1-.88623^2)
sd1=sqrt(var4)
sd1
```

### c.

```{r}
fctn4<-function(x){(1/2)*x*exp((-x^2)/4)}
prob5=integrate(fctn4,lower=mean4-(2*sd1),upper=mean4+(2*sd1))
prob5

mean4-(2*sd1)
mean4+(2*sd1)
```

### d.

```{r}
prob6=integrate(fctn4,lower=0, upper=6)
prob6

1-.9998766
```

It's not very likely.

## Question 5.

### a.

```{r}
mean5=2/(2+9)
mean5
var5=(2*9)/(((2+9)^2)*(2+9+1))
var5
```

### b.

```{r}
fctn5 <- function(x){x*(1-x)^8}
integrate(fctn5,lower=0, upper=1)
fctn6 <- function(x){(x*(1-x)^8)/((0.01111111))}
prob7=integrate(fctn6,lower=0,upper=.40)
prob7

1-.9536427
```

### c.

```{r}
prob8=integrate(fctn6,lower=0,upper=.10)
prob8
```

## Question 6 (MS 5.114 - Page 232)

### a.

```{r}

```

alpha = 2 and beta = 16

### b.

```{r}
mean5=((16)^(1/2))*.88623#table 6 appendix B
mean5
var5=(16)*(1-.88623^2)#table 6 appendix B
var5
```

### c.

```{r}
fctn7<-function(x){(1/8)*x*exp((-x^2)/16)}
prob9=integrate(fctn7,lower=0,upper=6)
prob9
1-0.8946008
```

## Question 7 (MS 6.2 - Page 239)

### a.

```{r}

```

$$p(x,y)=\frac{1}{36},1\leq x\leq36,1\leq y \leq 36$$

### b.

```{r}

```

$$
p_1(x)=\sum_{x=0}^{6}\frac{1}{36}=\frac{6}{36}=\frac{1}{6}, 1\leq x \leq 36\\
p_2(y)=\sum_{y=0}^{6}\frac{1}{36}=\frac{6}{36}=\frac{1}{6}, 1\leq y \leq 36
$$

### c.

```{r}

```

$$
p_1(x|y)=\frac{p(x, y)}{p_2(y)}=\frac{\frac{1}{36}}{\frac{1}{6}}=\frac{1}{6}\\
p_2(x|y)=\frac{p(x, y)}{p_1(y)}=\frac{\frac{1}{36}}{\frac{1}{6}}=\frac{1}{6}
$$

### d.

```{r}

```

$$
\bullet \;p_1(x)=p_1(x|y)\;and\;p_2(y)=p_2(y|x)\\
\bullet \; p(x, y)=p_1(x)p_2(y)
$$

Due to the fulfillment of the listed conditions, x and y are independent.

## Question 8

### a.

```{r}

```

$$
p(1,1)=\frac{1}{7}\\
p(1,2)=0\\
p(1,3)=0\\
\\------------\\
p(2,1)=\frac{2}{7}\\
p(2,2)=0\\
p(2,3)=0\\
\\------------\\
p(3,1)=\frac{1}{7}\\
p(3,2)=\frac{2}{7}\\
p(3,3)=\frac{1}{7}\\
$$

### b.

```{r}

```

$$
p_1(x)\\
p_1(1)=\frac{1}{7}\\
p_1(2)=\frac{2}{7}\\
p_1(3)=\frac{4}{7}
$$

### c.

```{r}

```

$$
p_2(y)\\
p_2(1)=\frac{4}{7}\\
p_2(2)=\frac{2}{7}\\
p_2(3)=\frac{1}{7}
$$

### d.

```{r}

```

$$
p_2(y|x)=\frac{p(x,y)}{p_1(x)}
\\------------\\
p(1|1)=1\\
p(1|2)=1\\
p(1|3)=\frac{1}{4}\\
\\------------\\
p(2|1)=0\\
p(2|2)=0\\
p(2|3)=\frac{1}{2}\\
\\------------\\
p(3,1)=0\\
p(3,2)=0\\
p(3,3)=\frac{1}{4}\\
$$

## Question 9. (MS 6.12 - Page 244)

### a.

```{r}

```

$$
f(y)=\int_{y}^{2y}f(x,y)dx=\int_{y}^{2y}\frac{e^{\frac{-y}{10}}}{10y}dx=\frac{e^{\frac{-y}{10}}}{10y}*(2y-y)=\frac{e^{\frac{-y}{10}}}{10}
$$

### b.

```{r}

```


$$
\lambda=\frac{1}{10} \implies E(Y)=\frac{1}{\lambda}=10
$$

## Question 10 (MS 6.14 - Page 245)

### a.

```{r}

```

$$
\int_{0}^{\infty}\int_{0}^{x}ce^{-x^2}dydx=\int_{0}^{\infty}xce^{-x^2}dx=\int_{0}^{\infty}\frac{c}{2}e^{-t}dx=\frac{c}{2}=1\\\implies\\c=2
$$

### b.

```{r}

```

$$
f_1(x)=\int_{0}^{x}2e^{-x^2}dydx=2xe^{-x^2} \implies \int_{-\infty}^{\infty}f_1(x)dx=1
$$

### c.

```{r}

```

$$
f_2(y|x)=\frac{f(x,y)}{f_1(x)}=\frac{2e^{-x^2}}{x2e^{-x^2}}=\frac{1}{x}
$$

## Question 11 (MS 6.51 - Page 253)

```{r}

```

$$
E(X)=\sum x P(X=x)=0\\
E(Y)=\sum y P(Y=y)=0\\
E(XY)=\sum xy P(X=x, Y=y)=0\\ \implies\\
COV(X,Y)=E(XY)-E(X)E(Y)=0
$$
but we also have

$$
P(X=x, Y=y) \neq P(X=x)*P(Y=y)\\
[ex:P(X=-1, Y=-1) \neq P(X=-1)*P(Y=-1) \iff\frac{1}{12} \neq \frac{1}{9})]
$$

for each x and y.

This means X and Y are not independent.

## Question 12 (MS 6.74 - Page 269)

### a.

$$
E(\bar Y)=\mu=\frac{b+a}{2}=\frac{3+1}{2}=2
$$
There are, on average, 2 parts per million uranium in each reservoir.

### b.

$$
Var(\bar Y)=\sigma ^2=\frac{\frac{(b-a)^2}{12}}{n}=\frac{\frac{(3-1)^2}{12}}{60}=\frac{\frac{1}{3}}{60}=\frac{1}{180}
$$

### c.

By The Central Limit Theorem, Y-bar has a Normal Distribution.

### d.

$$
P(1.5 \leq \bar Y \leq2.5)=P(\bar Y \leq 2.5) - P(\bar Y \leq 1.5)\\=P(\frac{2.5-2}{\sqrt{\frac{1}{180}}})-P(\frac{1.5-2}{\sqrt{\frac{1}{180}}})\\=P(Z \leq6.71)-P(Z\leq-6.71)\\1-0=1
$$

### e.

$$
P(\bar Y > 2.2)=1-P(\bar Y \leq 2.2)\\
=1-P(\frac{2.2-2}{\sqrt{\frac{1}{180}}})\\
=1-P(Z\leq2.68)\\
=1-.9963=.0037
$$

## Question 13 (MS 6.90 - Page 273)

### a.

$$
P(Z < \frac{(1+.5)-20*.4}{\sqrt{20*.4*.6}})=.0015
$$

### b.

$$
P(Z > \frac{(11-.5)-20*.4}{\sqrt{20*.4*.6}})=.1271
$$

### c.

$$
P(X<2)=.0005\\
P(X>10)=.1275
$$

The normal approximation did a good job for the second one but pretty bad on the first.









