---
title: "lab14"
author: "Harley Combest"
date: "`r Sys.Date()`"
output:
  html_document:
      toc: yes
      toc_float: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks


## Task 1

```{r}
getwd()
```

## Task 2

```{r}
mylsq=function(x,y){
ssxx=sum((x-mean(x))^2 )
ssxy=sum(  (x-mean(x))*(y-mean(y)) ) ## fill in the missing portion
b1hat=ssxy/ssxx
b0hat=(mean(y)- (b1hat*mean(x)) )   ##  fill in the missing portion

plot1=plot(x=x,y=y)

abline(coef=c(b0hat,b1hat),lwd=2,col="blue")

return(list(b0hat=b0hat,b1hat=b1hat )) #fill in the missing portion
}

x=1:20
set.seed(29);y=4+6*x + rnorm(20,0,5)
mylsq(x=x,y=y)
slr=lm(y~x);summary(slr)
```

## Task 3

```{r}
mypred=function(x,b0,b1){
ym=b0+b1+mean(x)
ym
}

xnew=15.5
b0=3.102214
b1=6.130737

ym=mypred(x=xnew,b0=b0,b1=1)
plot(x=x,y=y)
abline(coef=c(b0,b1),lwd=2,col="blue")
points(xnew,ym,cex=2,pch=1,col="blue")
```

```{r}
sweetness=read.csv("OJUICE.csv")
x1=sweetness$Pectin
y1=sweetness$SweetIndex
mylsq(x=x1,y=y1)
```


a) Least squares regression line: 6.252068 - 0.002310626x.

b) These variables represent the set decrease of sweetness via pectin.

c) 

```{r}
6.252068 - 0.002310626 * (300)
```


a)

```{r}
drillrock=read.csv("DRILLROCK.csv")
plot(drillrock)
```

b)

```{r}
lm(drillrock)
```

```{r}
x2=drillrock$DEPTH
y2=drillrock$TIME
scattergram1=mylsq(x=x2,y=y2)
scattergram1
```

Least Squares Prediction: 4.789603 + 0.01438785x.

c)

```{r}
plot(x=x2,y=y2)
abline(coef = (c(4.789603,0.01438785)))
```

d)

These variables represent the fixed cost of drilling.

## Task 4

```{r}
mysq=function(x,y){
n=length(x) # or y
ssxx=sum((x-mean(x))^2 )
ssxy=sum((x-mean(x))*(y-mean(y))) ## fill in the missing portion
b1hat=ssxy/ssxx
b0hat=mean(y)- (b1hat * mean(x))  ##  fill 
yhat=b0hat+(b1hat*mean(x))  ##  fill 
ssr=sum((y-yhat)^2)
sq=(ssr)/(n-2)##  fill 
return(list(ssr=ssr,sq=sq))
}

mysq(x=x,y=y)
var(y)

5.543077-5.196635
```

Pretty close. 0.346442 close.

a)

```{r}
carbon=read.csv("CARBON.csv")
x3=carbon$PilotPlant
y3=carbon$LabFurnace
scattergram2=plot(x=x3,y=y3)
```

b)

```{r}
mylsq(x=x3,y=y3)
```

These parameters represent the determined rise of carbon content in a lab furnace given the carbon content in a pilot plant.

c)

```{r}
mysq(x=x3,y=y3)
```

d)

```{r}
sqrt(0.2477217)
```

There is a 0.4977165 carbon content average difference between lab furnace carbon content from the mean lab furnace carbon content based on pilot plant carbon content.