---
title: "lab8"
author: "Harley Combest"
date: "`r Sys.Date()`"
output:
  html_document:
      toc: yes
      toc_float: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks


```{r}
(2*2)*(10*.4*.6)+(3*3)*(10*.4*.6)+(1)*(10*.4*.6)
(2^2)*(10*.4*.6)+(3^2)*(10*.4*.6)+(10*.4*.6)
pnorm(7, 10, sd=3) - pnorm(3, 10, 3)
pbinom(q=6, 10, prob=.4)-pbinom(2, 10, .4)
1 - pnorm(7,mean=5,sd=sqrt(10*.5*.5))
```



## Task 1

```{r}
getwd()
```

## Task 2

```{r}
unif1=runif(10,0,5)
unif1
mean1=(0+5)/2
mean1
var1=(5-0)/2
var1
mean(unif1)
var(unif1)
```

The sample mean is generally smaller and the sample variation much smaller than their population counterparts.

```{r}
T=sum(unif1)
T
yBar=T/10
yBar
```

```{r}
myclt=function(n,iter,a=0,b=5){
y=runif(n*iter,a,b)#A
data=matrix(y,nr=n,nc=iter,byrow=TRUE)#B
sm=apply(data,2,sum)#C
h=hist(sm,plot=FALSE)
hist(sm,col=rainbow(length(h$mids)),freq=FALSE,main="Distribution of the sum of uniforms")
curve(dnorm(x,mean=n*(a+b)/2,sd=sqrt(n*(b-a)^2/12)),add=TRUE,lwd=2,col="Blue")
sm
}
w=myclt(n=50,iter=10000,a=5,b=10)#D
```

A creates a uniform distribution with lower limit 0 and upper limit 5 with a sample size of n*iter.

B creates a matrix of n rows and iter columns.

C creates an array of values from the matrix created in B.

D calls on the function with input values n=50, iter=10000, a=5, and b=10.

```{r}
mean(w)
var(w)
```

```{r}
myclt2=function(n,iter,a=0,b=5){
y=runif(n*iter,a,b)#A
data=matrix(y,nr=n,nc=iter,byrow=TRUE)#B
sm=apply(data,2,mean)#C
h=hist(sm,plot=FALSE)
hist(sm,col=rainbow(length(h$mids)),freq=FALSE,main="Distribution of the sample means")
curve(dnorm(x,mean=(a+b)/2,sd=(b-a)/(sqrt(12*n))),add=TRUE,col="Red",lty=2,lwd=3)
sm
}
w1=myclt2(n=50,iter=10000,a=5,b=10)#D
mean(w1)
var(w1)
```

## Task 3

```{r}
mycltu=function(n,iter,a=0,b=10){
## r-random sample from the uniform
y=runif(n*iter,a,b)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax
## Now we can make the histogram
hist(w,freq=FALSE,  ylim=c(0,ymax), main=paste("Histogram of sample mean",
"\n", "sample size= ",n,sep=""),xlab="Sample mean")
## add a density curve made from the sample distribution
lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=(a+b)/2,sd=(b-a)/(sqrt(12*n))),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
## Add the density from which the samples were taken
curve(dunif(x,a,b),add=TRUE,lwd=4)

}
mycltu(n=20,iter=100000)
```

It uses the 2 for representing the 2 matrix columns in an array.

It takes 2000000 entries from the matrix.

That's the formula for the standard deviation of a uniform distribution.

```{r}
mycltu(n=1,iter=10000)
mycltu(n=2,iter=10000)
mycltu(n=3,iter=10000)
mycltu(n=5,iter=10000)
mycltu(n=10,iter=10000)
mycltu(n=30,iter=10000)
```

The function's distribution closely resembles a normal distribution in each instance save for the extremes of samples size; i.e. 1 and 30 produce the most offnorm distributions.

## Task 4

```{r}
mycltb=function(n,iter,p=0.5,...){

## r-random sample from the Binomial
y=rbinom(n*iter,size=n,prob=p)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax

## Now we can make the histogram
## freq=FALSE means take a density
hist(w,freq=FALSE,  ylim=c(0,ymax),
main=paste("Histogram of sample mean","\n", "sample size= ",n,sep=""),
xlab="Sample mean",...)
## add a density curve made from the sample distribution
#lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=n*p,sd=sqrt(p*(1-p))),add=TRUE,col="Red",lty=2,lwd=3) 

}


mycltb(n=5,iter=10000,p=0.5)

```

```{r}
mycltb(n=4,iter=10000,p=0.3)
mycltb(n=5,iter=10000,p=0.3)
mycltb(n=10,iter=10000,p=0.3)
mycltb(n=20,iter=10000,p=0.3)
```

Greater sample sizes yield distributions more fine tuned to the normal curve.

```{r}
mycltb(n=4,iter=10000,p=0.7)
mycltb(n=5,iter=10000,p=0.7)
mycltb(n=10,iter=10000,p=0.7)
mycltb(n=20,iter=10000,p=0.7)
```

The same trend as before exists for this p.

```{r}
mycltb(n=4,iter=10000,p=0.5)
mycltb(n=5,iter=10000,p=0.5)
mycltb(n=10,iter=10000,p=0.5)
mycltb(n=20,iter=10000,p=0.5)
```

I conclude that p does not affect the "normalness" of this distribution so much as sample size.

## Task 5

```{r}
mycltp=function(n,iter,lambda=10,...){

## r-random sample from the Poisson
y=rpois(n*iter,lambda=lambda)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax

## Make a suitable layout for graphing
layout(matrix(c(1,1,2,3),nr=2,nc=2, byrow=TRUE))

## Now we can make the histogram
hist(w,freq=FALSE,  ylim=c(0,ymax), col=rainbow(max(w)),
main=paste("Histogram of sample mean","\n", "sample size= ",n," iter=",iter," lambda=",lambda,sep=""),
xlab="Sample mean",...)
## add a density curve made from the sample distribution
#lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=lambda,sd=sqrt(lambda/n)),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve

# Now make a new plot
# Since y is discrete we should use a barplot
barplot(table(y)/(n*iter),col=rainbow(max(y)), main="Barplot of sampled y", ylab ="Rel. Freq",xlab="y" )
x=0:max(y)
plot(x,dpois(x,lambda=lambda),type="h",lwd=5,col=rainbow(max(y)),
main="Probability function for Poisson", ylab="Probability",xlab="y")
}


mycltp(n=2,iter=10000,lambda=4)
mycltp(n=3,iter=10000,lambda=4)
mycltp(n=5,iter=10000,lambda=4)
mycltp(n=10,iter=10000,lambda=4)
mycltp(n=20,iter=10000,lambda=4)

mycltp(n=2,iter=10000,lambda=10)
mycltp(n=3,iter=10000,lambda=10)
mycltp(n=5,iter=10000,lambda=10)
mycltp(n=10,iter=10000,lambda=10)
mycltp(n=20,iter=10000,lambda=10)
```

## Task 6

```{r}
myPackage1::studentCltp(n=20,iter=10000,lambda=10)
```

