---
title: "Lab2"
author: "Harley Combest"
date: "`r Sys.Date()`"
output: 
    html_document:
      toc: yes
      toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

```{r}
getwd()
```

## Task 2

```{r}
epagas=read.csv("EPAGAS.csv")
head(epagas)
```

## Task 3

```{r}
mpg=epagas$MPG

z=(mpg-mean(mpg))/(sd(mpg))
mean(z) #pretty close to zero
sd(z)^2

print("------------------")

#any mpg with an absolute z-score greater than 3 is an outlier.
z[abs(z)>3]
mpg[abs(z)>3]

print("------------------")

#possible outliers
z[abs(z)>=2 & abs(z)<=3]
mpg[abs(z)>=2 & abs(z)<=3]

#df1=data.frame(mpgs=mpg, zscores=z)
#df1[order(df1$mpgs, decreasing=FALSE),]

library(lattice)
mycol = ifelse(abs(z)>3, "Red",
        ifelse(abs(z)>=2 &abs(z)<=3,"Blue", "Black"))
dotplot(mpg, col=mycol)
```

## Task 4

```{r}
boxplot(mpg, data=epagas, main="Box plot for MPG", col="black", notch=TRUE, horizontal=TRUE)
```

By Chebyshev's Theorem 75% of the data within 2 standard deviations of the mean of the mpg data.

```{r}
boundedAbove=mean(mpg)+(2*sd(mpg))
boundedBelow=mean(mpg)-(2*sd(mpg))

numberOfMpgs=length(mpg)

epagasInInterval=with(epagas, epagas[MPG >= boundedBelow & MPG <= boundedAbove,])

numberOfMpgsInInterval=length(epagasInInterval)

percentInInterval=numberOfMpgsInInterval/numberOfMpgs
percentInInterval
```

Chebyshev's Theorem matches up with the data.

According to The Empirical Rule, approximately 95% of the data lies within 2 standard deviations of the mean.

This corresponds well.

The empirical rule is valid in this case because the distribution of the data is roughly bell shaped in its relative frequencies.
